
# About the project

*Write a short description about the course and add a link to your GitHub repository here. This is an R Markdown (.Rmd) file so you should use R Markdown syntax.*

```{r}
# This is a so-called "R chunk" where you can write R code.

date()

```

###Its great to learn of R tools which is very useful in the Research and analysis sector 
The text continues here.


### 2.1.2 Reading in the Global Burden of Disease example dataset

```{r}
library(tidyverse)
gbd_short <- read_csv("https://raw.githubusercontent.com/KimmoVehkalahti/RHDS/master/data/global_burden_disease_cause-year.csv")
View(gbd_short)
```

## 2.2 Variable types and why we care

**Note:** Some outputs may differ a bit from the book, because R is constantly developed further. One example is the following. (We have modified the code chunk a bit to reveal the outputs shown in the book.)

```{r}
library(tidyverse)
typesdata <- read_csv("https://raw.githubusercontent.com/KimmoVehkalahti/RHDS/master/data/typesdata.csv")
spec(typesdata)
typesdata
```

```{r}
typesdata_faulty <- read_csv("https://raw.githubusercontent.com/KimmoVehkalahti/RHDS/master/data/typesdata_faulty.csv")
spec(typesdata_faulty)
typesdata_faulty
```

### 2.2.1 Numeric variables (continuous)

```{r}
typesdata$measurement %>% mean()
measurement_mean <- typesdata$measurement %>% mean()

measurement_mean == 3.333333

(0.10 + 0.05) == 0.15

library(tidyverse)
near(0.10+0.05, 0.15)

near(measurement_mean, 3.333333, 0.000001)
```

### 2.2.2 Character variables

```{r}
library(tidyverse)
typesdata %>%
  count(group)

typesdata %>%
  count(group, sort = TRUE)

# all ids are unique:
typesdata %>% 
  count(id, sort = TRUE)

# we add in a duplicate row where id = ID3,
# then count again:
typesdata %>% 
  add_row(id = "ID3") %>% 
  count(id, sort = TRUE)
```

### 2.2.3 Factor variables (categorical)

(just read - more about this topic will follow later...)

### 2.2.4 Date/time variables

Remember to install the lubridate package first...

```{r}
library(lubridate) # install.packages("lubridate") # makes working with dates easier
current_datetime <- Sys.time()
current_datetime

my_datetime <- "2020-12-01 12:00"
my_datetime

# error that is explained in the book... (you may put '#' in front of it)
my_datetime - current_datetime

current_datetime %>% class()
my_datetime %>% class()
```

2.2.4 continued...

```{r}
my_datetime_converted <- ymd_hm(my_datetime)
my_datetime_converted

# now it will work:
my_datetime_converted - current_datetime

my_datesdiff <- my_datetime_converted - current_datetime
my_datesdiff %>% class()

ymd_hm("2021-01-02 12:00") + my_datesdiff

# another error challenge here... (again you may put '#' in front of that afterwards)
560/my_datesdiff
560/as.numeric(my_datesdiff)
```

2.2.4 continued...

```{r}
parse_date_time("12:34 07/Jan'20", "%H:%M %d/%b'%y")

Sys.time()
Sys.time() %>% format("%H:%M on %B-%d (%Y)")

Sys.time() %>% format("Happy days, the current time is %H:%M %B-%d (%Y)!")
```

## 2.3 Objects and functions

```{r}
library(tidyverse)
mydata <- tibble(
  id   = 1:4,
  sex  = c("Male", "Female", "Female", "Male"),
  var1 = c(4, 1, 2, 3),
  var2 = c(NA, 4, 5, NA),
  var3 = c(2, 1, NA, NA)
)
```

### 2.3.1 data frame/tibble

### 2.3.2 Naming objects

```{r}
mydata
```

### 2.3.3 Function and its arguments

```{r}
mydata$var1

mean(mydata$var1)

mean(mydata$var2)

mean(mydata$sex)

mean(mydata$var2, na.rm = TRUE)

Sys.time()
```

### 2.3.4 Working with objects

```{r}
a <- 103

a

seq(15, 30)

example_sequence <- seq(15, 30)

example_sequence <- example_sequence/2

example_sequence
```

### 2.3.5 \<- and =

```{r}
mean_result <- mean(mydata$var2, na.rm = TRUE)
```

### 2.3.6 Recap: object, function, input, argument

## 2.4 Pipe - %\>%

```{r}
library(tidyverse)
mydata$var1 %>% mean()

mean_result <- mydata$var1 %>% mean()
```

### 2.4.1 Using . to direct the pipe

```{r}
mydata %>% 
  lm(var1~var2, data = .)
```

## 2.5 Operators for filtering data

```{r}
gbd_short %>% 
  filter(year < 1995)

gbd_short %>% 
  filter(year <= 1995)

gbd_short %>% 
  filter(year == 1995)
```

2.5 continued...

```{r}
# do you see what is wrong here? (you may fix it or hide it with '#' afterwards)
gbd_short %>% 
 filter(year = 1995)

gbd_short %>% 
  filter(year == 1995 | year == 2017)

gbd_short %>% 
  filter(year == max(year) | year == min(year))
```

### 2.5.1 Worked examples

```{r}
mydata_year2000 <- gbd_short %>% 
  filter(year == 2000)

mydata_year2000
```

2.5.1 continued...

```{r}
new_data_selection <- gbd_short %>% 
  filter((year == 1990 | year == 2013) & cause == "Communicable diseases")

new_data_selection
```

2.5.1 continued...

```{r}
# Or we can get rid of the extra brackets around the years
# by moving cause into a new filter on a new line:

new_data_selection <- gbd_short %>% 
  filter(year == 1990 | year == 2013) %>% 
  filter(cause == "Communicable diseases")

new_data_selection
```

2.5.1 continued...

```{r}
  
# Or even better, we can include both in one filter() call, as all
# separate conditions are by default joined with "&":

new_data_selection <- gbd_short %>% 
  filter(year == 1990 | year == 2013,
         cause == "Communicable diseases")

new_data_selection
```

## 2.6 The combine function: c()

```{r}
gbd_short$cause %>% unique()

gbd_short %>% 
  # also filtering for a single year to keep the result concise
  filter(year == 1990) %>% 
  filter(cause == "Communicable diseases" | cause == "Non-communicable diseases")

gbd_short %>% 
  filter(year == 1990) %>% 
  filter(cause %in% c("Communicable diseases", "Non-communicable diseases"))
```

## 2.7 Missing values (NAs) and filters

```{r}
mydata

mydata %>% 
  filter(is.na(var2))

mydata %>% 
  filter(!is.na(var2))

mydata %>% 
  filter(var2 != 5)

mydata %>% 
  filter(var2 != 5 | is.na(var2))
```

2.7 continued...

```{r}
subset1 <- mydata %>% 
  filter(var2 == 5)

subset2 <- mydata %>% 
  filter(! var2 == 5)

subset1; subset2

nrow(mydata)
nrow(subset1)
nrow(subset2)

nrow(subset1) + nrow(subset2) == nrow(mydata)

subset3 <- mydata %>% 
  filter(is.na(var2))

nrow(subset1) + nrow(subset2) + nrow(subset3) == nrow(mydata)
```

## 2.8 Creating new columns - mutate()

```{r}
typesdata

typesdata$measurement

typesdata$measurement/2

typesdata %>% 
  mutate(measurement/2)

typesdata %>% 
  mutate(measurement_half = measurement/2)
```

2.8 continued...

```{r}
mydata$`Nasty column name`

# or

mydata %>% 
 select(`Nasty column name`)
```

2.8 continued...

```{r}
typesdata_modified <- typesdata %>% 
  mutate(measurement_half = measurement/2)

typesdata_modified
```

2.8 continued...

```{r}
library(lubridate)
typesdata %>% 
  mutate(reference_date   = ymd_hm("2020-01-01 12:00"),
         dates_difference = reference_date - date) %>% 
  select(date, reference_date, dates_difference)

typesdata %>% 
  mutate(mean_measurement = mean(measurement))

typesdata %>% 
  mutate(mean_measurement     = mean(measurement)) %>% 
  mutate(measurement_relative = measurement/mean_measurement) %>% 
  select(matches("measurement"))

```

### 2.8.1 Worked example/exercise

```{r}
typesdata %>% 
  mutate(reference_date   = ymd_hm("2020-01-01 12:00"),
         dates_difference = reference_date - date) %>% 
  mutate(dates_difference = round(dates_difference)) %>% 
  select(matches("date"))
```

## 2.9 Conditional calculations - if_else()

```{r}
typesdata %>% 
  mutate(above_threshold = if_else(measurement > 3,
                                   "Above three",
                                   "Below three"))
```

## 2.10 Create labels - paste()

```{r}
typesdata %>% 
  mutate(plot_label = paste(id,
                            "was last measured at", date,
                            ", and the value was",    measurement)) %>% 
  select(plot_label)

pastedata <- tibble(year  = c(2007, 2008, 2009),
                   month = c("Jan", "Feb", "March"),
                   day   = c(1, 2, 3))
pastedata
```

2.10 continued...

```{r}
pastedata %>% 
  mutate(date = paste(day, month, year, sep = "-"))

library(lubridate)

pastedata %>% 
  mutate(date = paste(day, month, year, sep = "-")) %>% 
  mutate(date = dmy(date))
```

## 2.11 Joining multiple datasets

```{r}
library(tidyverse)
patientdata <- read_csv("https://raw.githubusercontent.com/KimmoVehkalahti/RHDS/master/data/patient_data.csv")
patientdata

labsdata <- read_csv("https://raw.githubusercontent.com/KimmoVehkalahti/RHDS/master/data/labs_data.csv")
labsdata

 full_join(patientdata, labsdata)

inner_join(patientdata, labsdata)

 left_join(patientdata, labsdata)

right_join(patientdata, labsdata)
```

### 2.11.1 Further notes about joins

```{r}
patientdata_new <- read_csv("https://raw.githubusercontent.com/KimmoVehkalahti/RHDS/master/data/patient_data_updated.csv")
patientdata_new

bind_rows(patientdata, patientdata_new)

labsdata_updated <- labsdata %>% 
  add_row(id = 5, measurement = 2.49)
labsdata_updated

left_join(patientdata, labsdata_updated)
```

Well done! That was active reading of Chapter 2.

Working will continue below with Chapter 3...

------------------------------------------------------------------------

# Chapter 3: Summarising data

Continue reading at <https://argoshare.is.ed.ac.uk/healthyr_book/summarising-data.html> and working with the R code chunks.

## 3.1 Get the data

```{r}
library(tidyverse)
gbd_full <- read_csv("https://raw.githubusercontent.com/KimmoVehkalahti/RHDS/master/data/global_burden_disease_cause-year-sex-income.csv")

# Creating a single-year tibble for printing and simple examples:
gbd2017 <- gbd_full %>% 
  filter(year == 2017)

View(gbd2017)
```

## 3.2 Plot the data

```{r}
gbd2017 %>% 
  # without the mutate(... = fct_relevel()) 
  # the panels get ordered alphabetically
  mutate(income = fct_relevel(income,
                              "Low",
                              "Lower-Middle",
                              "Upper-Middle",
                              "High")) %>% 
  # defining the variables using ggplot(aes(...)):
  ggplot(aes(x = sex, y = deaths_millions, fill = cause)) +
  # type of geom to be used: column (that's a type of barplot):
  geom_col(position = "dodge") +
  # facets for the income groups:
  facet_wrap(~income, ncol = 4) +
  # move the legend to the top of the plot (default is "right"):
  theme(legend.position = "top")

```

## 3.3 Aggregating: group_by(), summarise()

```{r}
gbd2017$deaths_millions %>% sum()

gbd2017 %>% 
  summarise(sum(deaths_millions))

gbd2017 %>% 
  group_by(cause) %>% 
  summarise(sum(deaths_millions))

gbd2017 %>% 
  group_by(cause, sex) %>% 
  summarise(sum(deaths_millions))
```

## 3.4 Add new columns: mutate()

```{r}
gbd2017 %>% 
  group_by(cause, sex) %>% 
  summarise(deaths_per_group = sum(deaths_millions)) %>% 
  ungroup() %>% 
  mutate(deaths_total = sum(deaths_per_group))
```

### 3.4.1 Percentages formatting: percent()

```{r}
# percent() function for formatting percentages come from library(scales)
library(scales)
gbd2017_summarised <- gbd2017 %>% 
  group_by(cause, sex) %>% 
  summarise(deaths_per_group = sum(deaths_millions)) %>% 
  ungroup() %>% 
  mutate(deaths_total    = sum(deaths_per_group),
         deaths_relative = percent(deaths_per_group/deaths_total))
gbd2017_summarised

# using values from the first row as an example:
round(100*4.91/55.74, 1) %>% paste0("%")

gbd2017_summarised %>% 
  mutate(deaths_relative = deaths_per_group/deaths_total)
```

## 3.5 summarise() vs mutate()

```{r}
gbd_summarised <- gbd2017 %>% 
  group_by(cause, sex) %>% 
  summarise(deaths_per_group = sum(deaths_millions)) %>% 
  arrange(sex)

gbd_summarised

gbd_summarised_sex <- gbd_summarised %>% 
  group_by(sex) %>% 
  summarise(deaths_per_sex = sum(deaths_per_group))

gbd_summarised_sex
```

3.5 continued...

```{r}
full_join(gbd_summarised, gbd_summarised_sex)

gbd_summarised %>% 
  group_by(sex) %>% 
  mutate(deaths_per_sex = sum(deaths_per_group))

gbd2017 %>% 
  group_by(cause, sex) %>% 
  summarise(deaths_per_group = sum(deaths_millions)) %>% 
  group_by(sex) %>% 
  mutate(deaths_per_sex  = sum(deaths_per_group),
         sex_cause_perc = percent(deaths_per_group/deaths_per_sex)) %>% 
  arrange(sex, deaths_per_group)
```

## 3.6 Common arithmetic functions - sum(), mean(), median(), etc.

```{r}
mynumbers <- c(1, 2, NA)
sum(mynumbers)

sum(mynumbers, na.rm = TRUE)
```

## 3.7 select() columns

```{r}
gbd_2rows <- gbd_full %>% 
  slice(1:2)

gbd_2rows

gbd_2rows %>% 
  select(cause, deaths_millions)

gbd_2rows %>% 
  select(cause, deaths = deaths_millions)

gbd_2rows %>% 
  rename(deaths = deaths_millions)

gbd_2rows %>% 
  select(year, sex, income, cause, deaths_millions)

gbd_2rows %>% 
  select(year, sex, everything())

gbd_2rows %>% 
  select(starts_with("deaths"))
```

## 3.8 Reshaping data - long vs wide format

```{r}
gbd_wide <- read_csv("https://raw.githubusercontent.com/KimmoVehkalahti/RHDS/master/data/global_burden_disease_wide-format.csv")
gbd_long <- read_csv("https://raw.githubusercontent.com/KimmoVehkalahti/RHDS/master/data/global_burden_disease_cause-year-sex.csv")

gbd_wide
gbd_long
```

### 3.8.1 Pivot values from rows into columns (wider)

```{r}
gbd_long %>% 
  pivot_wider(names_from = year, values_from = deaths_millions)

gbd_long %>% 
  pivot_wider(names_from = sex, values_from = deaths_millions) %>% 
  mutate(Male - Female)

gbd_long %>% 
  pivot_wider(names_from = c(sex, year), values_from = deaths_millions)
```

### 3.8.2 Pivot values from columns to rows (longer)

```{r}
gbd_wide %>% 
  pivot_longer(matches("Female|Male"), 
               names_to = "sex_year", 
               values_to = "deaths_millions") %>% 
  slice(1:6)

gbd_wide %>% 
  select(matches("Female|Male"))
```

### 3.8.3 separate() a column into multiple columns

```{r}
gbd_wide %>% 
  # same pivot_longer as before
  pivot_longer(matches("Female|Male"), 
               names_to = "sex_year", 
               values_to = "deaths_millions") %>% 
  separate(sex_year, into = c("sex", "year"), sep = "_", convert = TRUE)
```

## 3.9 arrange() rows

```{r}
gbd_long %>% 
  arrange(deaths_millions) %>% 
  # first 3 rows just for printing:
  slice(1:3)

gbd_long %>% 
  arrange(-deaths_millions) %>% 
  slice(1:3)

gbd_long %>% 
  arrange(desc(sex)) %>% 
  # printing rows 1, 2, 11, and 12
  slice(1,2, 11, 12)
```

### 3.9.1 Factor levels

```{r}
gbd_factored <- gbd_long %>% 
  mutate(cause = factor(cause))

gbd_factored$cause %>% levels()

gbd_factored <- gbd_factored %>% 
  mutate(cause = cause %>% 
           fct_relevel("Injuries"))

gbd_factored$cause %>% levels()
```

### 3.10.1 Exercise - pivot_wider()

Look at Table 3.4 on page <https://argoshare.is.ed.ac.uk/healthyr_book/exercises.html>

```{r}
gbd_long <- read_csv("https://raw.githubusercontent.com/KimmoVehkalahti/RHDS/master/data/global_burden_disease_cause-year-sex.csv")

# Solution:

gbd_long %>% 
  pivot_wider(names_from = cause, values_from = deaths_millions)
```

### 3.10.2 Exercise - group_by(), summarise()

```{r}
gbd_full <- read_csv("https://raw.githubusercontent.com/KimmoVehkalahti/RHDS/master/data/global_burden_disease_cause-year-sex-income.csv")

glimpse(gbd_full)

summary_data1 <- 
  gbd_full %>% 
  group_by(year) %>% 
  summarise(total_per_year = sum(deaths_millions))

summary_data1

summary_data2 <- 
  gbd_full %>% 
  group_by(year, cause) %>% 
  summarise(total_per_cause = sum(deaths_millions))

summary_data2
```

### 3.10.3 Exercise - full_join(), percent()

```{r}
# Solution:

library(scales)
full_join(summary_data1, summary_data2) %>% 
  mutate(percentage = percent(total_per_cause/total_per_year)) 
```

### 3.10.4 Exercise - mutate(), summarise()

```{r}
# Solution:

gbd_full %>% 
  # aggregate to deaths per cause per year using summarise()
  group_by(year, cause) %>% 
  summarise(total_per_cause = sum(deaths_millions)) %>% 
  # then add a column of yearly totals using mutate()
  group_by(year) %>% 
  mutate(total_per_year = sum(total_per_cause)) %>% 
  # add the percentage column
  mutate(percentage = percent(total_per_cause/total_per_year)) %>% 
  # select the final variables for better viewing
  select(year, cause, percentage) %>% 
  pivot_wider(names_from = cause, values_from = percentage)
```

### 3.10.5 Exercise - filter(), summarise(), pivot_wider()

```{r}
# Solution:

gbd_full %>% 
  filter(year == 1990) %>% 
  group_by(income, sex) %>% 
  summarise(total_deaths = sum(deaths_millions)) %>% 
  pivot_wider(names_from = income, values_from = total_deaths)
```

Wow! That was all for Chapter 3. GOOD JOB.

The next chapter will take us in plotting awesome graphs. Let's proceed!

------------------------------------------------------------------------

# Chapter 4: Different types of plots

Look at Figure 4.1 at <https://argoshare.is.ed.ac.uk/healthyr_book/chap04-h1.html>.

You will now re-create the figure step by step!

## 4.1 Get the data

```{r}
library(tidyverse)
library(gapminder) # install.packages("gapminder")

glimpse(gapminder)

gapminder$year %>% unique()
gapminder$country %>% n_distinct()
gapminder$continent %>% unique()

gapdata2007 <- gapminder %>% 
  filter(year == 2007)

gapdata2007

# loads the gapminder dataset from the package environment
# into your Global Environment
gapdata <- gapminder
```

## 4.2 Anatomy of ggplot explained

```{r}
# recommended form:
gapdata2007 %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp))

# NOT recommended form:
ggplot(gapdata2007, aes(x = gdpPercap, y = lifeExp))

# just a schematic example of using the pipe:
#
# data %>% 
#   filter(...) %>% 
#   mutate(...) %>% 
#   ggplot(aes(...)) +
#   ...
```

4.2 continued...

```{r}
gapdata2007 %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point()

gapdata2007 %>% 
  ggplot(aes(x = continent, y = lifeExp)) +
  geom_point()

gapdata2007 %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp, colour = continent)) +
  geom_point()

gapdata2007 %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp, colour = continent)) +
  geom_point(shape = 1)

```

4.2 continued...

```{r}
gapdata2007 %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp, colour = continent)) +
  geom_point(shape = 1) +
  facet_wrap(~continent)

gapdata2007 %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp, colour = continent)) +
  geom_point(shape = 1) +
  facet_wrap(~pop > 50000000)

gapdata2007 %>% 
  ggplot(aes(x = gdpPercap/1000, y = lifeExp, colour = continent)) +
  geom_point(shape = 1) +
  facet_wrap(~continent) +
  theme_bw()
```

## 4.3 Set your theme - grey vs white

```{r}
theme_set(theme_bw())

library(tidyverse)
theme_set(theme_bw())
```

## 4.4 Scatter plots/bubble plots

```{r}
gapdata2007 %>% 
  ggplot(aes(x = gdpPercap/1000, y = lifeExp, size = pop)) +
  geom_point()

gapdata2007 %>% 
  ggplot(aes(x = gdpPercap/1000, y = lifeExp, size = pop)) +
  geom_point(shape = 1, alpha = 0.5)
```

## 4.5 Line plots/time series plots

```{r}
gapdata %>% 
  filter(country == "United Kingdom") %>% 
  ggplot(aes(x = year, y = lifeExp)) +
  geom_line()

gapdata %>% 
  ggplot(aes(x = year, y = lifeExp)) +
  geom_line()

gapdata %>% 
  ggplot(aes(x = year, y = lifeExp, group = country)) +
  geom_line()

```

### 4.5.1 Exercise

Look at Figure 4.9 on page <https://argoshare.is.ed.ac.uk/healthyr_book/line-plotstime-series-plots.html>

```{r}
gapdata %>% 
  ggplot(aes(x = year, y = lifeExp, group = country)) +
  geom_line()
```

Try to transform the above graph following these instructions:

-   Colour lines by continents: add colour = continent inside aes();
-   Continents on separate facets: + facet_wrap(\~continent);
-   Use a nicer colour scheme: + scale_colour_brewer(palette = "Paired")

## 4.6 Bar plots

### 4.6.1 Summarised data

```{r}
gapdata2007 %>% 
  filter(country %in% c("United Kingdom", "France", "Germany")) %>% 
  ggplot(aes(x = country, y = lifeExp)) +
  geom_col() 
```

### 4.6.2 Countable data

```{r}
gapdata2007 %>% 
  count(continent)

gapdata2007 %>% 
  ggplot(aes(x = continent)) +
  geom_bar()

gapdata2007 %>% 
  ggplot(aes(x = continent, colour = country)) +
  geom_bar(fill = NA) +
  theme(legend.position = "none")
```

### 4.6.3 colour vs fill

### 4.6.4 Proportions

```{r}
gapdata2007 %>% 
  ggplot(aes(x = "Global", fill = continent)) + 
  geom_bar()
```

### 4.6.5 Exercise

Look at Figure 4.13 on the page <https://argoshare.is.ed.ac.uk/healthyr_book/bar-plots.html> and try to recreate it!

```{r}

```

## 4.7 Histograms

```{r}
gapdata2007 %>% 
  ggplot(aes(x = lifeExp)) +
  geom_histogram(binwidth = 10)
```

## 4.8 Box plots

```{r}
gapdata2007 %>% 
  ggplot(aes(x = continent, y = lifeExp)) +
  geom_boxplot()
```

## 4.9 Multiple geoms, multiple aes()

```{r}
# (1)
gapdata2007 %>% 
  ggplot(aes(x = continent, y = lifeExp)) +
  geom_boxplot() +
  geom_point()

# (2)
gapdata2007 %>% 
  ggplot(aes(x = continent, y = lifeExp)) +
  geom_boxplot() +
  geom_jitter()
```

4.9 continued...

```{r}
# (3)
gapdata2007 %>% 
  ggplot(aes(x = continent, y = lifeExp, colour = continent)) +
  geom_boxplot() +
  geom_jitter()

# (4)
gapdata2007 %>% 
  ggplot(aes(x = continent, y = lifeExp)) +
  geom_boxplot() +
  geom_jitter(aes(colour = continent))
```

### 4.9.1 Worked example - three geoms together

```{r}
label_data <- gapdata2007 %>% 
  group_by(continent) %>% 
  filter(lifeExp == max(lifeExp)) %>% 
  select(country, continent, lifeExp)

# since we filtered for lifeExp == max(lifeExp)
# these are the maximum life expectancy countries at each continent:
label_data

gapdata2007 %>% 
  ggplot(aes(x = continent, y = lifeExp)) +
  # First geom - boxplot
  geom_boxplot() +
  # Second geom - jitter with its own aes(colour = )
  geom_jitter(aes(colour = continent)) +
  # Third geom - label, with its own dataset (label_data) and aes(label = )
  geom_label(data = label_data, aes(label = country))
```

Try the suggested experiments given on page <https://argoshare.is.ed.ac.uk/healthyr_book/multiple-geoms-multiple-aes.html> with the R code above. (Copy-paste the above code chunk below to experiment with it!)

## 4.10 All other types of plots

## 4.11 Solutions

You will find solutions to Exercises 4.5.1 and 4.6.5 on the page <https://argoshare.is.ed.ac.uk/healthyr_book/solutions.html>

## 4.12 Extra: Advanced examples

(this is all extra material, if you are curious and have some time to check it!)

```{r}
gapdata %>% 
  filter(continent == "Europe") %>% 
  ggplot(aes(y      = fct_reorder(country, lifeExp, .fun=max),
             x      = lifeExp,
             colour = year)) +
  geom_point(shape = 15, size = 2) +
  scale_colour_distiller(palette = "Greens", direction = 1) +
  theme_bw()
```

another example:

```{r}
gapdata2007 %>% 
  group_by(continent) %>% 
  mutate(country_number = seq_along(country)) %>% 
  ggplot(aes(x = continent)) +
  geom_bar(aes(colour = continent), fill = NA, show.legend = FALSE) +
  geom_text(aes(y = country_number, label = country), vjust = 1)+
  geom_label(aes(label = continent), y = -1) +
  theme_void()
```

Great! That was it for Chapter 4.

The next chapter will take us further in fine tuning plots. It can be well considered EXTRA MATERIAL, so if you are curious, just go on and try it, too!

**You may also skip Chapter 5 and proceed straight to Chapter 6.**

You may come back in Chapter 5 anytime later, to learn the fine-tuning tricks!
